---
title: "Untitled"
author: "Andy White"
date: "November 17, 2018"
output:
  word_document: default
  html_document: default
---

##Step	1:	Load	the	data
We	will	use	the	airquality	data	set,	which you	should	already	have	as	part	of	your	R	installation.
```{r part1}
#install.packages("reshape2")
library(reshape2)
#install.packages("tidyr")
library(tidyr)
#install.packages("ggplot2")
library(ggplot2)
#install.packages("dplyr")
library(dplyr)

arqual <- airquality
#arqual

```

##Step	2:	Clean	the	data
After	you	load	the	data,	there	will	be	some	NAs	in	the	data.	You	need	to	figure	out	what	to	do	about	those	nasty	NAs.


``` {r removenas}
#I could remove the NA's with Na.omit. However this will omit a lot of values
arqual2 <- na.omit(airquality)

#This also will omit NAs
complete.cases(arqual)
which(!complete.cases(arqual))
na_vec <- which(!complete.cases(arqual)) #this will give us the na rows that are present 
#arqual[-na_vec,]

#Or replace NA's with the mean:

colSums(is.na(arqual))

meanOZ <- mean(arqual$Ozone,na.rm = TRUE)
meanOZ
meanSolar <- mean(arqual$Solar.R,na.rm = TRUE)
meanSolar


Ozone <- replace_na(arqual$Ozone,meanOZ)
arqual$Ozone2 <- replace_na(arqual$Ozone,meanOZ)
arqual$Solar2 <- replace_na(arqual$Solar.R,meanSolar)
#arqual

```


##Step	3:	Understand	the	data	distribution
Create	the	following	visualizations using	ggplot:
. Histograms	for	each	of	the	variables		
. Boxplot	for	Ozone
. Boxplot for	wind	values	(round	the	wind	to	get	a	good	number	of	"buckets")

```{r step3}
hist1 <- ggplot(arqual,aes(x=Ozone2,fill="white"))+geom_histogram(binwidth = 20)
#updated binwidth to be less granular however givent the 37 values at the "mean" 
hist1
#will choose the data set that we used omit.na
hist2 <-ggplot(arqual2,aes(x=Ozone,fill="white"))+geom_histogram(bins=10)
hist2
hist_sol1 <- ggplot(arqual,aes(x=Solar2,fill="white"))+geom_histogram(binwidth = 30)
hist_sol1

hist_sol2 <-ggplot(arqual2,aes(x=Solar.R,fill="white"))+geom_histogram(bins=10)
hist_sol2

ggplot(arqual,aes(x=Wind,color="white"))+geom_histogram()
ggplot(arqual,aes(x=Temp,color="white"))+geom_histogram()
ggplot(arqual,aes(x=Month,color="white"))+geom_histogram()
ggplot(arqual,aes(x=Day,color="white"))+geom_histogram()

ggplot(arqual,aes(x=factor(0),y=Ozone2))+geom_boxplot()
#group by month:
ggplot(arqual,aes(group=Month,x=Month,y=Ozone2))+geom_boxplot()

ggplot(arqual,aes(group=Month,x=Month,y=Wind))+geom_boxplot()


```


##Step	3:	Explore	how	the	data	changes	over	time
First,	make	sure	to	create	appropriate	dates	(this	data	was	from	1973).	Then	create	line	charts for	ozone,	temp,	wind	and	solar.R (one	line	chart	for	each,	and	then	one	chart	with	4	lines,	each	having	a	different	color).
Create	these	visualizations	using	ggplot. Note	that	for	the	chart	with	4	lines,	you	need	to	think	about	how	to	effectively	use	the	yaxis.

``` {r step4}
arqual3 <- arqual
arqual3$Date <- gsub(" ","",paste(arqual$Month,"/",arqual$Day,"/",1973))
 #the function above concatenates the data into a look alike date value
arqual3$Date <-as.Date(arqual3$Date,"%m/%d/%Y")                        

arqual2$Date <-gsub(" ","",paste(arqual2$Month,"/",arqual2$Day,"/",1973))
arqual2$Date <- as.Date(arqual2$Date,"%m/%d/%Y")
#this will convert the new date column to date values

#Using the omited data as the the means show too many flat spaces for ozone
a <- ggplot(arqual2,aes(x=Date))+geom_line(aes(y=Ozone))
b <- ggplot(arqual2,aes(x=Date))+geom_line(aes(y=Solar.R))
c <- ggplot(arqual2,aes(x=Date))+geom_line(aes(y=Wind))
d <- ggplot(arqual2,aes(x=Date))+geom_line(aes(y=Temp))

ggplot(arqual2,aes(x=Date))+geom_line(aes(y=Ozone))


ggplot(arqual2,aes(x=Date))+geom_line(aes(y=Ozone),colour="red")+geom_line(aes(y=Temp),colour="green")+geom_line(aes(y=Wind),colour="yellow") + geom_line(aes(y=Solar.R),colour="black")

```


##Step	4:	Look	at	all	the data	via	a	Heatmap
Create a	heatmap,	with	each	day	along	the	x-axis	and	ozone,	temp,	wind	and	solar.r	along	the	y-axis,	and	days	as	rows	along	the	y-axis.	 Creat	the	heatmap	using	geom_tile (this defines	the	ggplot geometry	to	be	'tiles'	as	opposed	to	'lines'	and	the	other	geometry	we	have	previously	used).
Note	that	you	need	to	figure	out	how	to	show	the	relative	change	equally	across	all	the	
variables.

##This question is kind of confusing confusing. I don't know what it is actually asking me to do. I have attempted what I think it is asking. 
```{r step5}


arqual4 <- arqual2 
#arqual4$daymon <- paste(arqual4$Day,arqual4$Month)
arqual4 <-arqual4[,-5:-6]

arqual4 <- gather(arqual4, ColNam, Data,1:4)
arqual4$percent_chg <- arqual4$Data/lag(arqual4$Data)-1
#this adds a column and uses the "Data" and then divides by previous line item
arqual4$percent_chg <- replace_na(arqual4$percent_chg,0)
#replace first value with 0
arqual4

#x=2
#ifelse()
#arqual4$percent_chg2<-while (x!=nrow(arqual4)) {
#  ifelse(arqual4[x,1]==arqual4[x-1,1],arqual4[x,4],arqual4[x,4]==0)
#  x<-x+1
#}
View(arqual4)
arqual4

ggplot(arqual4,aes(x=Date,percent_chg))+geom_tile(aes(y=factor(ColNam)))

ggplot(arqual4,aes(x=Date,y=factor(ColNam)))+geom_tile(aes(fill=Data)) 

#ggplot(arqual2,aes(x=Date,y=mp))+geom_tile()+scale_fill_gradient(colours=colorRampPalette(c('blue','red'))(100))


```


##Step	5:	Look	at	all	the	data	via	a	scatter	chart
Create	a	scatter	chart (using	ggplot	geom_point),	with	the	x-axis	representing	the	wind,	the	
y-axis	representing	the	temperature,	the	size	of	each	dot	representing	the	ozone	and	the	
color	representing	the	solar.R
```{r Step6}
plot_graph <- ggplot(arqual2,aes(x=Wind,y=Temp))
plot_graph <- plot_graph  + geom_point(aes(size=Ozone,color = Solar.R))
plot_graph
```
##Step	6:	Final	Analysis
. Do	you	see	any	patterns	after	exploring	the	data?	
 I can definitely see that there is useful patterns in the last data, the higher temp and with less wind, showed higher Ozone. This makes intuitive sense, since ozone alerts happen on hotter days with less wind, with larger amounts of solar. 
. What	was	the	most	useful	visualization?
I liked the last one, the other ones were difficult to read and didn't break up the data enough. 









